---
title: "Express Scan - association plaque characteristics and OLINK plasma proteins"
author:
- Joost M. Mekke
- MD
output:
  html_notebook:
    cache: yes
    code_folding: hide
    collapse: yes
    df_print: paged
    fig.align: center
    fig_caption: yes
    fig_height: 10
    fig_retina: 2
    fig_width: 12
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  html_document:
    df_print: paged
    toc: yes
mainfont: Helvetica
subtitle: An 'Express Scan' project
editor_options:
  chunk_output_type: inline
---
```{r global_options, include = FALSE}
# further define some knitr-options.
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, fig.path = 'Figures/', eval=TRUE, warning=FALSE, message = FALSE)
```

# Preparation
Clean the environment
```{r ClearEnvironment, include = FALSE}
rm(list=ls())
```

Set locations, and the working directory.
```{r LocalSystem, include=FALSE}
### MacBook Pro
PROJECT_loc = "C:\\Users\\anton\\Documents\\UMCU\\Joost_data"
AEDB_loc = "C:\\Users\\anton\\Documents\\UMCU\\Joost_data"

### SOME VARIABLES WE NEED DOWN THE LINE
PROJECTNAME = "OLINK_SlideToolkit"

cat("\nCreate a new analysis directory...\n")
ifelse(!dir.exists(file.path(PROJECT_loc, "\\",PROJECTNAME)), 
       dir.create(file.path(PROJECT_loc, "\\",PROJECTNAME)), 
       FALSE)
ANALYSIS_loc = paste0(PROJECT_loc,"/",PROJECTNAME)
# Plots
ifelse(!dir.exists(file.path(ANALYSIS_loc, "/PLOTS")), 
       dir.create(file.path(ANALYSIS_loc, "/PLOTS")), 
       FALSE)
PLOT_loc = paste0(ANALYSIS_loc,"/PLOTS")
# Quality Control plots
ifelse(!dir.exists(file.path(PLOT_loc, "/QC")), 
       dir.create(file.path(PLOT_loc, "/QC")), 
       FALSE)
QC_loc = paste0(PLOT_loc,"/QC")
# Output files
ifelse(!dir.exists(file.path(ANALYSIS_loc, "/OUTPUT")), 
       dir.create(file.path(ANALYSIS_loc, "/OUTPUT")), 
       FALSE)
OUT_loc = paste0(ANALYSIS_loc, "/OUTPUT")
# COX analysis
ifelse(!dir.exists(file.path(ANALYSIS_loc, "/COX")), 
       dir.create(file.path(ANALYSIS_loc, "/COX")), 
       FALSE)
COX_loc = paste0(ANALYSIS_loc, "/COX")
# Baseline characteristics
ifelse(!dir.exists(file.path(ANALYSIS_loc, "/BASELINE")), 
       dir.create(file.path(ANALYSIS_loc, "/BASELINE")), 
       FALSE)
BASELINE_loc = paste0(ANALYSIS_loc, "/BASELINE")

cat("\nSetting working directory and listing its contents.\n")
setwd(paste0(PROJECT_loc))
getwd()
list.files()
```
A package-installation function.
```{r Function: installations, include = FALSE}
install.packages.auto <- function(x) { 
  x <- as.character(substitute(x)) 
  if(isTRUE(x %in% .packages(all.available = TRUE))) { 
    eval(parse(text = sprintf("require(\"%s\")", x)))
  } else { 
    # Update installed packages - this may mean a full upgrade of R, which in turn
    # may not be warrented. 
    # update.install.packages.auto(ask = FALSE) 
    eval(parse(text = sprintf("install.packages(\"%s\", dependencies = TRUE, repos = \"https://cloud.r-project.org/\")", x)))
  }
  if(isTRUE(x %in% .packages(all.available = TRUE))) { 
    eval(parse(text = sprintf("require(\"%s\")", x)))
  } else {
    if (!requireNamespace("BiocManager"))
      install.packages("BiocManager")
    # BiocManager::install() # this would entail updating installed packages, which in turned may not be warrented
    eval(parse(text = sprintf("BiocManager::install(\"%s\")", x)))
    eval(parse(text = sprintf("require(\"%s\")", x)))
  }
}
```
Load those packages.
```{r Setting: loading_packages, include = FALSE}
install.packages.auto("readr")
install.packages.auto("optparse")
install.packages.auto("tools")
install.packages.auto("dplyr")
install.packages.auto("tidyr")
install.packages.auto("naniar")

# To get 'data.table' with 'fwrite' to be able to directly write gzipped-files
# Ref: https://stackoverflow.com/questions/42788401/is-possible-to-use-fwrite-from-data-table-with-gzfile
# install.packages("data.table", repos = "https://Rdatatable.gitlab.io/data.table")
library(data.table)

install.packages.auto("tidyverse")
install.packages.auto("knitr")
install.packages.auto("DT")
install.packages.auto("MASS")
# install.packages.auto("Seurat") # latest version

# Install the devtools package from Hadley Wickham
install.packages.auto('devtools')

install.packages.auto("haven")
install.packages.auto("sjlabelled")
install.packages.auto("sjPlot")
install.packages.auto("labelled")
install.packages.auto("tableone")
install.packages.auto("foreign")
install.packages.auto("ggpubr")
install.packages.auto("ggplot2")
install.packages.auto("survival")
install.packages.auto("finalfit")
install.packages.auto("car")

install.packages.auto("factoextra")
```
We will create a datestamp and define the Utrecht Science Park Colour Scheme.
```{R Setting: Colors, include = FALSE}

Today = format(as.Date(as.POSIXlt(Sys.time())), "%Y%m%d")
Today.Report = format(as.Date(as.POSIXlt(Sys.time())), "%A, %B %d, %Y")

### UtrechtScienceParkColoursScheme
###
### WebsitetoconvertHEXtoRGB:http://hex.colorrrs.com.
### Forsomefunctionsyoushoulddividethesenumbersby255.
### 
###	No.	Color			      HEX	(RGB)						              CHR		  MAF/INFO
###---------------------------------------------------------------------------------------
###	1	  yellow			    #FBB820 (251,184,32)				      =>	1		or 1.0>INFO
###	2	  gold			      #F59D10 (245,157,16)				      =>	2		
###	3	  salmon			    #E55738 (229,87,56)				      =>	3		or 0.05<MAF<0.2 or 0.4<INFO<0.6
###	4	  darkpink		    #DB003F ((219,0,63)				      =>	4		
###	5	  lightpink		    #E35493 (227,84,147)				      =>	5		or 0.8<INFO<1.0
###	6	  pink			      #D5267B (213,38,123)				      =>	6		
###	7	  hardpink		    #CC0071 (204,0,113)				      =>	7		
###	8	  lightpurple	    #A8448A (168,68,138)				      =>	8		
###	9	  purple			    #9A3480 (154,52,128)				      =>	9		
###	10	lavendel		    #8D5B9A (141,91,154)				      =>	10		
###	11	bluepurple		  #705296 (112,82,150)				      =>	11		
###	12	purpleblue		  #686AA9 (104,106,169)			      =>	12		
###	13	lightpurpleblue	#6173AD (97,115,173/101,120,180)	=>	13		
###	14	seablue			    #4C81BF (76,129,191)				      =>	14		
###	15	skyblue			    #2F8BC9 (47,139,201)				      =>	15		
###	16	azurblue		    #1290D9 (18,144,217)				      =>	16		or 0.01<MAF<0.05 or 0.2<INFO<0.4
###	17	lightazurblue	  #1396D8 (19,150,216)				      =>	17		
###	18	greenblue		    #15A6C1 (21,166,193)				      =>	18		
###	19	seaweedgreen	  #5EB17F (94,177,127)				      =>	19		
###	20	yellowgreen		  #86B833 (134,184,51)				      =>	20		
###	21	lightmossgreen	#C5D220 (197,210,32)				      =>	21		
###	22	mossgreen		    #9FC228 (159,194,40)				      =>	22		or MAF>0.20 or 0.6<INFO<0.8
###	23	lightgreen	  	#78B113 (120,177,19)				      =>	23/X
###	24	green			      #49A01D (73,160,29)				      =>	24/Y
###	25	grey			      #595A5C (89,90,92)				        =>	25/XY	or MAF<0.01 or 0.0<INFO<0.2
###	26	lightgrey		    #A2A3A4	(162,163,164)			      =>	26/MT
###
###	ADDITIONAL COLORS
###	27	midgrey			#D7D8D7
###	28	verylightgrey	#ECECEC"
###	29	white			#FFFFFF
###	30	black			#000000
###----------------------------------------------------------------------------------------------

uithof_color = c("#FBB820","#F59D10","#E55738","#DB003F","#E35493","#D5267B",
                 "#CC0071","#A8448A","#9A3480","#8D5B9A","#705296","#686AA9",
                 "#6173AD","#4C81BF","#2F8BC9","#1290D9","#1396D8","#15A6C1",
                 "#5EB17F","#86B833","#C5D220","#9FC228","#78B113","#49A01D",
                 "#595A5C","#A2A3A4", "#D7D8D7", "#ECECEC", "#FFFFFF", "#000000")

uithof_color_legend = c("#FBB820", "#F59D10", "#E55738", "#DB003F", "#E35493",
                        "#D5267B", "#CC0071", "#A8448A", "#9A3480", "#8D5B9A",
                        "#705296", "#686AA9", "#6173AD", "#4C81BF", "#2F8BC9",
                        "#1290D9", "#1396D8", "#15A6C1", "#5EB17F", "#86B833",
                        "#C5D220", "#9FC228", "#78B113", "#49A01D", "#595A5C",
                        "#A2A3A4", "#D7D8D7", "#ECECEC", "#FFFFFF", "#000000")

#ggplot2 default color palette
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]}
```
## Plasma protein data
Loading Athero-Express plasma protein measurements from 2019/2020 as measured using OLINK.
```{r LoadAE PlasmaProteins OLINK}
library(openxlsx)
AEDB_PlasmaProtein_OLINK_CVD2raw <- openxlsx::read.xlsx(paste0(AEDB_loc, "/20200706_AtheroExpress_OlinkData_forR.xlsx"), sheet = "CVD2_forR")
AEDB_PlasmaProtein_OLINK_CVD3raw <- openxlsx::read.xlsx(paste0(AEDB_loc, "/20200706_AtheroExpress_OlinkData_forR.xlsx"), sheet = "CVD3_forR")
AEDB_PlasmaProtein_OLINK_CMraw <- openxlsx::read.xlsx(paste0(AEDB_loc, "/20200706_AtheroExpress_OlinkData_forR.xlsx"), sheet = "CM_forR")
AEDB_PlasmaProtein_OLINK_ProteinInfo <- openxlsx::read.xlsx(paste0(AEDB_loc, "/20200706_AtheroExpress_OlinkData_forR.xlsx"), sheet = "ProteinInfo")

AEDB_PlasmaProtein_OLINK_CVD2 <- AEDB_PlasmaProtein_OLINK_CVD2raw %>% filter(QC_Warning_CVD2 == "Pass")
AEDB_PlasmaProtein_OLINK_CVD3 <- AEDB_PlasmaProtein_OLINK_CVD3raw %>% filter(QC_Warning_CVD3 == "Pass")
AEDB_PlasmaProtein_OLINK_CM <- AEDB_PlasmaProtein_OLINK_CMraw %>% filter(QC_Warning_CM == "Pass")

table(AEDB_PlasmaProtein_OLINK_CVD2raw$QC_Warning_CVD2)
table(AEDB_PlasmaProtein_OLINK_CVD2$QC_Warning_CVD2)

table(AEDB_PlasmaProtein_OLINK_CVD3raw$QC_Warning_CVD3)
table(AEDB_PlasmaProtein_OLINK_CVD3$QC_Warning_CVD3)

table(AEDB_PlasmaProtein_OLINK_CMraw$QC_Warning_CM)
table(AEDB_PlasmaProtein_OLINK_CM$QC_Warning_CM)

AEDB_PlasmaProtein_OLINK_CVD2$Plate_ID <- NULL
AEDB_PlasmaProtein_OLINK_CVD3$Plate_ID <- NULL
AEDB_PlasmaProtein_OLINK_CVD2$Order <- NULL
AEDB_PlasmaProtein_OLINK_CVD3$Order <- NULL
AEDB_PlasmaProtein_OLINK_CM$Order <- NULL

temp <- merge(AEDB_PlasmaProtein_OLINK_CVD2, AEDB_PlasmaProtein_OLINK_CVD3, by.x = "STUDY_NUMBER", by.y = "STUDY_NUMBER",
              sort = FALSE, all.x = TRUE)

AEDB_PlasmaProtein_OLINK <- merge(temp, AEDB_PlasmaProtein_OLINK_CM, by.x = "STUDY_NUMBER", by.y = "STUDY_NUMBER",
              sort = FALSE, all.x = TRUE)

AEDB_PlasmaProtein_OLINK$Plate_ID[AEDB_PlasmaProtein_OLINK$Plate_ID == "ERA_UMC_AE_cardiometabolic_plt1_29-10-19"] <- "plate 1"
AEDB_PlasmaProtein_OLINK$Plate_ID[AEDB_PlasmaProtein_OLINK$Plate_ID == "ERA_UMC_AE_Cardiometabolic_plt2"] <- "plate 2"
AEDB_PlasmaProtein_OLINK$Plate_ID[AEDB_PlasmaProtein_OLINK$Plate_ID == "ERA_UMC_AE_Cardiometabolic_plt3"] <- "plate 3"
AEDB_PlasmaProtein_OLINK$Plate_ID[AEDB_PlasmaProtein_OLINK$Plate_ID == "ERA_UMC_AE_Cardiometabolic_plt4"] <- "plate 4"
AEDB_PlasmaProtein_OLINK$Plate_ID[AEDB_PlasmaProtein_OLINK$Plate_ID == "ERA_UMC_AE_Cardiometabolic_plt5"] <- "plate 5"
AEDB_PlasmaProtein_OLINK$Plate_ID[AEDB_PlasmaProtein_OLINK$Plate_ID == "ERA_UMC_AE_Cardiometabolic_pl6"] <- "plate 6"
AEDB_PlasmaProtein_OLINK$Plate_ID[AEDB_PlasmaProtein_OLINK$Plate_ID == "SMART_CM_plt10"] <- "plate 10"
AEDB_PlasmaProtein_OLINK$Plate_ID[AEDB_PlasmaProtein_OLINK$Plate_ID == "SMART_plt11_CM"] <- "plate 11"

olink_proteins <- c("BMP6", "ANGPT1", "ADM", "CD40L", "SLAMF7", "PGF", "ADAMTS13", "BOC", "IL4RA", "SRC", "IL1ra", "IL6", "TNFRSF10A", "STK4", "IDUA", 
                    "TNFRSF11A", "PAR1", "TRAILR2", "PRSS27", "TIE2", "TF", "IL1RL2", "PDGF_subunit_B", "IL27", "IL17D", "CXCL1", "LOX1", "Gal9", "GIF", "SCF", 
                    "IL18", "FGF21", "PIgR", "RAGE", "SOD2", "CTRC", "FGF23", "SPON2", "GH", "FS", "GLO1", "CD84", "PAPPA", "SERPINA12", "REN", "DECR1", 
                    "MERTK", "KIM1", "THBS2", "TM", "VSIG2", "AMBP", "PRELP", "HO1", "XCL1", "IL16", "SORT1", "CEACAM8", "PTX3", "PSGL1", "CCL17", "CCL3", 
                    "MMP7", "IgG_Fc_receptor_IIb", "ITGB1BP2", "DCN", "Dkk1", "LPL", "PRSS8", "AGRP", "HBEGF", "GDF2", "FABP2", "THPO", "MARCO", "GT", "BNP", 
                    "MMP12", "ACE2", "PDL2", "CTSL1", "hOSCAR", "TNFRSF13B", "TGM2", "LEP", "CA5A", "HSP_27", "CD4", "NEMO", "VEGFD", "PARP1", "HAOX1", 
                    "TNFRSF14", "LDL_receptor", "ITGB2", "IL17RA", "TNFR2", "MMP9", "EPHB4", "IL2RA", "OPG", "ALCAM", "TFF3", "SELP", "CSTB", "MCP1", "CD163", 
                    "Gal3", "GRN", "NTproBNP", "BLM_hydrolase", "PLC", "LTBR", "Notch_3", "TIMP4", "CNTN1", "CDH5", "TLT2", "FABP4", "TFPI", "PAI", "CCL24", 
                    "TR", "TNFRSF10C", "GDF15", "SELE", "AZU1", "DLK1", "SPON1", "MPO", "CXCL16", "IL6RA", "RETN", "IGFBP1", "CHIT1", "TRAP", "GP6", "PSPD", 
                    "PI3", "EpCAM", "APN", "AXL", "IL1RT1", "MMP2", "FAS", "MB", "TNFSF13B", "PRTN3", "PCSK9", "UPAR", "OPN", "CTSD", "PGLYRP1", "CPA1", "JAMA", 
                    "Gal4", "IL1RT2", "SHPS1", "CCL15", "CASP3", "uPA", "CPB1", "CHI3L1", "ST2", "tPA", "SCGB3A2", "EGFR", "IGFBP7", "CD93", "IL18BP", "COL1A1", 
                    "PON3", "CTSZ", "MMP3", "RARRES2", "ICAM2", "KLK6", "PDGF_subunit_A", "TNFR1", "IGFBP2", "vWF", "PECAM1", "MEPE", "CCL16", "PRCP","CA1", 
                    "ICAM1", "CHL1", "TGFBI", "ENG", "PLTP", "SERPINA7", "IGFBP3", "CR2", "SERPINA5", "FCGR3B", "IGFBP6", "CDH1", "CCL5", "CCL14", "GNLY", 
                    "NOTCH1", "PAM", "PROC", "CST3", "NCAM1", "PCOLCE", "LILRB1", "MET", "LTBP2", "IL7R", "VCAM1", "SELL", "F11", "COMP", "CA4", "PTPRS", 
                    "MBL2", "TIMP1", "ANGPTL3", "REG3A", "SOD1", "CD46", "ITGAM", "TNC", "NID1", "CFHR5", "SPARCL1", "PLXNB2", "MEGF9", "ANG", "ST6GAL1", 
                    "DPP4", "REG1A", "QPCT", "FCN2", "FETUB", "CES1", "CRTAC1", "TCN2", "PRSS2", "ICAM3", "SAA4", "CNDP1", "FCGR2A", "NRP1", "EFEMP1", "TIMD4", 
                    "FAP", "TIE1", "THBS4", "F7", "GP1BA", "LYVE1", "CA3", "TGFBR3", "DEFA1", "CD59", "APOM", "OSMR", "LILRB2", "UMOD", "CCL18", "COL18A1", 
                    "LCN2", "KIT", "C1QTNF1", "AOC3", "GAS6", "IGLC2", "PLA2G7", "TNXB", "MFAP5", "VASN", "LILRB5", "C2")

length(olink_proteins)

olink_proteins_rank = unlist(lapply(olink_proteins, paste0, "_rankNorm"))

rm(temp)
```

### Inspect OLINK data

We know that the proteins are not normally distributed and therefore we will standardise them as follows: 

`z = ( x - μ ) / σ`

Where for each sample, `x` equals the value of the variable, `μ` (_mu_) equals the mean of `x`, and `σ` (_sigma_) equals the standard deviation of `x`.

```{r OLINKStandardize}
for(PROTEIN in 1:length(olink_proteins)){
  # AEDB_PlasmaProtein_OLINK$Z <- NULL
  var.temp.z = olink_proteins_rank[PROTEIN]
  var.temp = olink_proteins[PROTEIN]
  
  cat(paste0("\nSelecting ", var.temp, " and standardising: ", var.temp.z,".\n"))
  cat(paste0("* changing ", var.temp, " to numeric.\n"))

  AEDB_PlasmaProtein_OLINK <- AEDB_PlasmaProtein_OLINK %>%
    mutate_each_(funs(as.numeric), olink_proteins) 
  
  cat(paste0("* standardising ", var.temp, 
             " (mean: ",round(mean(!is.na(AEDB_PlasmaProtein_OLINK[,var.temp])), digits = 6),
             ", n = ",sum(!is.na(AEDB_PlasmaProtein_OLINK[,var.temp])),").\n"))
  
  AEDB_PlasmaProtein_OLINK <- AEDB_PlasmaProtein_OLINK %>%
      mutate_at(vars(var.temp), 
        list(Z = ~ (AEDB_PlasmaProtein_OLINK[,var.temp] - mean(AEDB_PlasmaProtein_OLINK[,var.temp], na.rm = TRUE))/sd(AEDB_PlasmaProtein_OLINK[,var.temp], na.rm = TRUE))
      )
  # str(AEDB_PlasmaProtein_OLINK$Z)
  cat(paste0("* renaming Z to ", var.temp.z,".\n"))
  AEDB_PlasmaProtein_OLINK[,var.temp.z] <- NULL
  names(AEDB_PlasmaProtein_OLINK)[names(AEDB_PlasmaProtein_OLINK) == "Z"] <- var.temp.z
}

rm(var.temp, var.temp.z)
```
Here we summarize some of these data in the subset of genetic data that passed QC.
```{r OLINKSummary}

for(PROTEIN in 1:length(olink_proteins)){
  # AEDB_PlasmaProtein_OLINK$Z <- NULL
  var.temp.z = olink_proteins_rank[PROTEIN]
  var.temp = olink_proteins[PROTEIN]
  
  cat(paste0("\nSummarising data for ",var.temp," [AU]; n = ",sum(!is.na(AEDB_PlasmaProtein_OLINK[,var.temp])),".\n"))
  print(summary(AEDB_PlasmaProtein_OLINK[,var.temp]))
  print(summary(AEDB_PlasmaProtein_OLINK[,var.temp.z]))

}
rm(var.temp, var.temp.z, PROTEIN)
```

```{r OLINKVisualize , include=FALSE}
require("ggpubr")
require("ggsci")

# mypal = pal_npg("nrc", alpha = 0.7)(9)
# mypal
# ## [1] "#E64B35B2" "#4DBBD5B2" "#00A087B2" "#3C5488B2" "#F39B7FB2" "#8491B4B2"
# ## [7] "#91D1C2B2" "#DC0000B2" "#7E6148B2"
# library("scales")
# show_col(mypal)

for(PROTEIN in 1:length(olink_proteins)){
  # metabolite_unit = ucorbioNMRDataDictionary[ucorbioNMRDataDictionary$Metabolite_NMR == NMRtargets[METABOLITE], "Concentration_NMR"]
  cat(paste0("\nProcessing metabolite [ ",olink_proteins[PROTEIN]," (AU)].\n"))

  var.temp = olink_proteins[PROTEIN]
  var.temp.z = paste0(olink_proteins[PROTEIN],"_rankNorm")
   
  dt.temp <- subset(AEDB_PlasmaProtein_OLINK, select = c("STUDY_NUMBER", var.temp, var.temp.z, "Plate_ID"))
  dt.temp[,2] <- as.numeric(dt.temp[,2])
  
  p1 <- ggpubr::gghistogram(dt.temp %>% filter(!is.na(Plate_ID)),
                    x = var.temp,
                    y = "..count..",
                    color = "#4DBBD5B2", fill = "#4DBBD5B2",
                    # palette = "npg",
                    rug = TRUE,
                    add = "mean",
                    xlab = paste0(var.temp," [AU]."),
                    ggtheme = theme_minimal())
  
  my_comparisons <- list( c("plate 1", "plate 2"), 
                          c("plate 1", "plate 3"), 
                          c("plate 1", "plate 4"), 
                          c("plate 1", "plate 5"), 
                          c("plate 1", "plate 6"), 
                          c("plate 1", "plate 10"), 
                          c("plate 1", "plate 11") )
  p2 <- ggpubr::ggboxplot(data = dt.temp %>% filter(!is.na(Plate_ID)),
                          x = "Plate_ID",
                          y = var.temp.z,
                          color = "Plate_ID",
                          palette = "npg",
                          add = c("mean", "jitter"),
                          # error.plot = "errorbar",
                          xlab = "plates used",
                          ylab = paste0(var.temp.z," [AU]."),
                          # ylim = c(0,4),
                          ggtheme = theme_minimal()) + stat_compare_means(method = "anova") + stat_compare_means(comparisons = my_comparisons) + stat_compare_means(label = "p.signif", method = "t.test", ref.group = "plate 1")
  
  
  p3 <- ggpubr::gghistogram(dt.temp %>% filter(!is.na(Plate_ID)),
                    x = var.temp.z,
                    y = "..count..",
                    color = "#91D1C2B2", fill = "#91D1C2B2",
                    # palette = "npg",
                    rug = TRUE,
                    add = "mean",
                    xlab = paste0(var.temp.z," [AU]."),
                    ggtheme = theme_minimal())
  
  require(patchwork)
  # p4 <- ((p1 / p3 ) | (p2))
  
  p4 <- ggpar(p1, legend = "" ) / ggpar(p2 + rotate_x_text(45), legend = "")  | ggpar(p3, legend = "right")
  
  print(p4)
  ggsave(filename = paste0(QC_loc, "/",Today,".",PROJECTNAME,".OLINK.",var.temp,".png"),
         plot = p4, device = "png", width = 20, height = 20)
  
}
  # rm(my_comparisons,
  #    p1, p2, p3, p4,
  #    var.temp, var.temp.z, dt.temp, PROTEIN)

```

```{r}
# Creating a new dataframe with only the normalized markers
AEDB_PlasmaProtein_OLINK_rank <- AEDB_PlasmaProtein_OLINK[,c(1,288:563)]
#AEDB_PlasmaProtein_OLINK_norm <- AEDB_PlasmaProtein_OLINK[,c(1:93,97:188,192:283)]
#write.csv2(AEDB_PlasmaProtein_OLINK_rank,file= paste0(OUT_loc, "/", Today, "OLINKplasmarank.csv"), sep=';', row.names=TRUE, quote=FALSE)

# Calculating what percent of the OLINK data per protein consists of missings
p_missing <- unlist(lapply(AEDB_PlasmaProtein_OLINK_rank, function(x) sum(is.na(x))))/nrow(AEDB_PlasmaProtein_OLINK_rank)
sort(p_missing[p_missing > 0], decreasing = TRUE)

# The following markers have over >70% missing value (which means that the measurements were under the LOD (limit of detection)): 
# -PRCP
# -ITGAM
# -SOD1
# -FAP
# -PARP1
# -REG3A
# -LCN2
# -LTBP2
# -QPCT
# -DEFA1
# -UMOD
# -PAPPA

x <- AEDB_PlasmaProtein_OLINK_rank[lapply(AEDB_PlasmaProtein_OLINK_rank, function(x) sum(is.na(x)) / length(x) ) < 0.70 ]
colnames(x)
# Check if the following OLINK proteins are indeed deleted from the dataframe
p_missing2 <- unlist(lapply(x, function(x) sum(is.na(x))))/nrow(x)
sort(p_missing2[p_missing2 > 0], decreasing = TRUE)

# Somehow STUDY_NUMBER is deleted from the dataframe, re-add STUDY_NUMBER so that merging of the dataframes is possible
# x$STUDY_NUMBER <- AEDB_PlasmaProtein_OLINK$STUDY_NUMBER

AEDB_PlasmaProtein_OLINK_rank <- x
colnames(AEDB_PlasmaProtein_OLINK_rank)
```

# Loading the ExpressScan data (this includes the AE clinical data as well)
```{r, include = FALSE}
#ExpressScan <-read.csv2(paste0(AEDB_loc,"/20210329ExpressScan.csv"))
ExpressScan <-read.csv2(paste0(AEDB_loc,"/20210518_ExpressScan_histology_data.csv"))
```

# Merging the OLINK data with the ExpressScan data
```{r, include = FALSE}
ExpressScan_OLINK <- merge(ExpressScan,AEDB_PlasmaProtein_OLINK_rank, by= "STUDY_NUMBER", all=FALSE)

#p_missing3 <- unlist(lapply(ExpressScan_OLINK, function(x) sum(is.na(x))))/nrow(ExpressScan_OLINK)
#sort(p_missing3[p_missing2 > 0], decreasing = TRUE)
```

```{r Analysis Functions, include = FALSE}
AEDB.CEA.samplesize = nrow(ExpressScan_OLINK)

# Function to grep data from glm()/lm()
GLM.CON <- function(fit, DATASET, x_name, y, verbose=c(TRUE,FALSE)){
  cat("Analyzing in dataset '", DATASET ,"' the association of '", x_name ,"' with '", y ,"' .\n")
  if (nrow(summary(fit)$coefficients) == 1) {
    output = c(DATASET, x_name, y, rep(NA,8))
    cat("Model not fitted; probably singular.\n")
  }else {
    cat("Collecting data.\n\n")
    effectsize = summary(fit)$coefficients[2,1]
    SE = summary(fit)$coefficients[2,2]
    OReffect = exp(summary(fit)$coefficients[2,1])
    CI_low = exp(effectsize - 1.96 * SE)
    CI_up = exp(effectsize + 1.96 * SE)
    tvalue = summary(fit)$coefficients[2,3]
    pvalue = summary(fit)$coefficients[2,4]
    R = summary(fit)$r.squared
    R.adj = summary(fit)$adj.r.squared
    sample_size = nrow(model.frame(fit))
    AE_N = AEDB.CEA.samplesize
    Perc_Miss = 100 - ((sample_size * 100)/AE_N)
    
    output = c(DATASET, x_name, y, effectsize, SE, OReffect, CI_low, CI_up, tvalue, pvalue, R, R.adj, AE_N, sample_size, Perc_Miss)
    
    if (verbose == TRUE) {
    cat("We have collected the following and summarize it in an object:\n")
    cat("Dataset...................:", DATASET, "\n")
    cat("Score/Exposure/biomarker..:", x_name, "\n")
    cat("Trait/outcome.............:", y, "\n")
    cat("Effect size...............:", round(effectsize, 6), "\n")
    cat("Standard error............:", round(SE, 6), "\n")
    cat("Odds ratio (effect size)..:", round(OReffect, 3), "\n")
    cat("Lower 95% CI..............:", round(CI_low, 3), "\n")
    cat("Upper 95% CI..............:", round(CI_up, 3), "\n")
    cat("T-value...................:", round(tvalue, 6), "\n")
    cat("P-value...................:", signif(pvalue, 8), "\n")
    print(R)
    cat("R^2.......................:", round(R, 6), "\n")
    cat("Adjusted r^2..............:", round(R.adj, 6), "\n")
    cat("Sample size of AE DB......:", AE_N, "\n")
    cat("Sample size of model......:", sample_size, "\n")
    cat("Missing data %............:", round(Perc_Miss, 6), "\n")
    } else {
      cat("Collecting data in summary object.\n")
    }
  }
  return(output)
  print(output)
}

```

```{r, include=FALSE}
# Choose one of the following plaque characteristics
# CD34: "Total_CD34_TissueCount_rank"
# CD66b: "Total_CD66b_TissueCount_rank"
# CD68: "Total_CD68_TissueCount_rank"
# SMA: "Total_SMA_TissueCount_rank"
# HE: "Total_HE_TissueCount_rank"
# EVG: "Total_EVG_Tissue_rank"
# GLYCC: "Total_GLYCC_Tissue_rank"

# And the correct accompanying plaque sizes
# CD34: PlaquesizeCD34_rank
# CD66b: "PlaquesizeCD66_rank"
# CD68: "PlaquesizeCD68_rank"
# SMA: "PlaquesizeSMA_rank"
# HE: "PlaquesizeHE_rank"
# EVG: "PlaquesizeEVG_rank"
# GLYCC:"PlaquesizeGlycc_rank"

# List of the 7 plaque characteristics
PLAQUE.CHARACTERISTICS.RANK = c("Total_CD34_TissueCount_rank", "Total_CD66b_TissueCount_rank", "Total_CD68_TissueCount_rank", "Total_SMA_TissueCount_rank", "Total_HE_TissueCount_rank", "Total_EVG_Tissue_rank", "Total_GLYCC_Tissue_rank") 

# List of the plaque size covariates, index matches that of PLAQUE.CHARACTERISTICS.RANK
COVARIATES_M1 = c("PlaquesizeCD34_rank", "PlaquesizeCD66b_rank", "PlaquesizeCD68_rank", "PlaquesizeSMA_rank", "PlaquesizeHE_rank", "PlaquesizeEVG_rank", "PlaquesizeGlycc_rank") 

olink_proteins_rank <- c("BMP6_rankNorm", "ANGPT1_rankNorm", "ADM_rankNorm", "CD40L_rankNorm", "SLAMF7_rankNorm", "PGF_rankNorm", "ADAMTS13_rankNorm", "BOC_rankNorm", "IL4RA_rankNorm", "SRC_rankNorm", "IL1ra_rankNorm", "IL6_rankNorm", "TNFRSF10A_rankNorm", "STK4_rankNorm", "IDUA_rankNorm", 
                    "TNFRSF11A_rankNorm", "PAR1_rankNorm", "TRAILR2_rankNorm", "PRSS27_rankNorm", "TIE2_rankNorm", "TF_rankNorm", "IL1RL2_rankNorm", "PDGF_subunit_B_rankNorm", "IL27_rankNorm", "IL17D_rankNorm", "CXCL1_rankNorm", "LOX1_rankNorm", "Gal9_rankNorm", "GIF_rankNorm", "SCF_rankNorm", 
                    "IL18_rankNorm", "FGF21_rankNorm", "PIgR_rankNorm", "RAGE_rankNorm", "SOD2_rankNorm", "CTRC_rankNorm", "FGF23_rankNorm", "SPON2_rankNorm", "GH_rankNorm", "FS_rankNorm", "GLO1_rankNorm", "CD84_rankNorm", "SERPINA12_rankNorm", "REN_rankNorm", "DECR1_rankNorm", 
                    "MERTK_rankNorm", "KIM1_rankNorm", "THBS2_rankNorm", "TM_rankNorm", "VSIG2_rankNorm", "AMBP_rankNorm", "PRELP_rankNorm", "HO1_rankNorm", "XCL1_rankNorm", "IL16_rankNorm", "SORT1_rankNorm", "CEACAM8_rankNorm", "PTX3_rankNorm", "PSGL1_rankNorm", "CCL17_rankNorm", "CCL3_rankNorm", 
                    "MMP7_rankNorm", "IgG_Fc_receptor_IIb_rankNorm", "ITGB1BP2_rankNorm", "DCN_rankNorm", "Dkk1_rankNorm", "LPL_rankNorm", "PRSS8_rankNorm", "AGRP_rankNorm", "HBEGF_rankNorm", "GDF2_rankNorm", "FABP2_rankNorm", "THPO_rankNorm", "MARCO_rankNorm", "GT_rankNorm", "BNP_rankNorm", 
                    "MMP12_rankNorm", "ACE2_rankNorm", "PDL2_rankNorm", "CTSL1_rankNorm", "hOSCAR_rankNorm", "TNFRSF13B_rankNorm", "TGM2_rankNorm", "LEP_rankNorm", "CA5A_rankNorm", "HSP_27_rankNorm", "CD4_rankNorm", "NEMO_rankNorm", "VEGFD_rankNorm", "HAOX1_rankNorm", 
                    "TNFRSF14_rankNorm", "LDL_receptor_rankNorm", "ITGB2_rankNorm", "IL17RA_rankNorm", "TNFR2_rankNorm", "MMP9_rankNorm", "EPHB4_rankNorm", "IL2RA_rankNorm", "OPG_rankNorm", "ALCAM_rankNorm", "TFF3_rankNorm", "SELP_rankNorm", "CSTB_rankNorm", "MCP1_rankNorm", "CD163_rankNorm", 
                    "Gal3_rankNorm", "GRN_rankNorm", "NTproBNP_rankNorm", "BLM_hydrolase_rankNorm", "PLC_rankNorm", "LTBR_rankNorm", "Notch_3_rankNorm", "TIMP4_rankNorm", "CNTN1_rankNorm", "CDH5_rankNorm", "TLT2_rankNorm", "FABP4_rankNorm", "TFPI_rankNorm", "PAI_rankNorm", "CCL24_rankNorm", 
                    "TR_rankNorm", "TNFRSF10C_rankNorm", "GDF15_rankNorm", "SELE_rankNorm", "AZU1_rankNorm", "DLK1_rankNorm", "SPON1_rankNorm", "MPO_rankNorm", "CXCL16_rankNorm", "IL6RA_rankNorm", "RETN_rankNorm", "IGFBP1_rankNorm", "CHIT1_rankNorm", "TRAP_rankNorm", "GP6_rankNorm", "PSPD_rankNorm", 
                    "PI3_rankNorm", "EpCAM_rankNorm", "APN_rankNorm", "AXL_rankNorm", "IL1RT1_rankNorm", "MMP2_rankNorm", "FAS_rankNorm", "MB_rankNorm", "TNFSF13B_rankNorm", "PRTN3_rankNorm", "PCSK9_rankNorm", "UPAR_rankNorm", "OPN_rankNorm", "CTSD_rankNorm", "PGLYRP1_rankNorm", "CPA1_rankNorm", "JAMA_rankNorm", 
                    "Gal4_rankNorm", "IL1RT2_rankNorm", "SHPS1_rankNorm", "CCL15_rankNorm", "CASP3_rankNorm", "uPA_rankNorm", "CPB1_rankNorm", "CHI3L1_rankNorm", "ST2_rankNorm", "tPA_rankNorm", "SCGB3A2_rankNorm", "EGFR_rankNorm", "IGFBP7_rankNorm", "CD93_rankNorm", "IL18BP_rankNorm", "COL1A1_rankNorm", 
                    "PON3_rankNorm", "CTSZ_rankNorm", "MMP3_rankNorm", "RARRES2_rankNorm", "ICAM2_rankNorm", "KLK6_rankNorm", "PDGF_subunit_A_rankNorm", "TNFR1_rankNorm", "IGFBP2_rankNorm", "vWF_rankNorm", "PECAM1_rankNorm", "MEPE_rankNorm", "CCL16_rankNorm","CA1_rankNorm", 
                    "ICAM1_rankNorm", "CHL1_rankNorm", "TGFBI_rankNorm", "ENG_rankNorm", "PLTP_rankNorm", "SERPINA7_rankNorm", "IGFBP3_rankNorm", "CR2_rankNorm", "SERPINA5_rankNorm", "FCGR3B_rankNorm", "IGFBP6_rankNorm", "CDH1_rankNorm", "CCL5_rankNorm", "CCL14_rankNorm", "GNLY_rankNorm", 
                    "NOTCH1_rankNorm", "PAM_rankNorm", "PROC_rankNorm", "CST3_rankNorm", "NCAM1_rankNorm", "PCOLCE_rankNorm", "LILRB1_rankNorm", "MET_rankNorm", "IL7R_rankNorm", "VCAM1_rankNorm", "SELL_rankNorm", "F11_rankNorm", "COMP_rankNorm", "CA4_rankNorm", "PTPRS_rankNorm", 
                    "MBL2_rankNorm", "TIMP1_rankNorm", "ANGPTL3_rankNorm", "CD46_rankNorm", "TNC_rankNorm", "NID1_rankNorm", "CFHR5_rankNorm", "SPARCL1_rankNorm", "PLXNB2_rankNorm", "MEGF9_rankNorm", "ANG_rankNorm", "ST6GAL1_rankNorm", 
                    "DPP4_rankNorm", "REG1A_rankNorm", "FCN2_rankNorm", "FETUB_rankNorm", "CES1_rankNorm", "CRTAC1_rankNorm", "TCN2_rankNorm", "PRSS2_rankNorm", "ICAM3_rankNorm", "SAA4_rankNorm", "CNDP1_rankNorm", "FCGR2A_rankNorm", "NRP1_rankNorm", "EFEMP1_rankNorm", "TIMD4_rankNorm", 
                    "TIE1_rankNorm", "THBS4_rankNorm", "F7_rankNorm", "GP1BA_rankNorm", "LYVE1_rankNorm", "CA3_rankNorm", "TGFBR3_rankNorm", "CD59_rankNorm", "APOM_rankNorm", "OSMR_rankNorm", "LILRB2_rankNorm", "CCL18_rankNorm", "COL18A1_rankNorm", 
                    "KIT_rankNorm", "C1QTNF1_rankNorm", "AOC3_rankNorm", "GAS6_rankNorm", "IGLC2_rankNorm", "PLA2G7_rankNorm", "TNXB_rankNorm", "MFAP5_rankNorm", "VASN_rankNorm", "LILRB5_rankNorm", "C2_rankNorm")

# Path to the file containing the risk scores (should be tab separated, have a single header line and contain all the risk scores and covariates that are to be used in the analysis)
score_file <- "C:\\Users\\anton\\Documents\\UMCU\\metagrs_comp_scores.txt"

# prs_methods contains a list of PRS methods as they occur in the score_file, not necessarily in the same order
#prs_methods <- c("CAD_PRSCS", "CAD_Rapido", "CAD_PLINK", "CAD_PRSice_B", "IS_PRSCS", "IS_Rapido", "IS_PLINK", "IS_PRSice_B", "CAD_PRSice_PF", "IS_PRSice_PF") # PRS methods of main project
prs_methods <- c("Inouye_CAD", "Inouye_CAD_pos", "Inouye_CAD_neg", "Abraham_IS", "Abraham_IS_pos", "Abraham_IS_neg") # For the metagrs comparison

scores <- read.table(score_file, sep="\t", header=TRUE)

GLM.results <- data.frame(matrix(NA, ncol = 15, nrow = 0))

### Loop used for plaque~prs associations
# for (ind.plaque in 1:length(PLAQUE.CHARACTERISTICS.RANK)) {
#   TRAIT = PLAQUE.CHARACTERISTICS.RANK[ind.plaque]
#   plaquesize_cov = COVARIATES_M1[ind.plaque]
# 
#   for (ind.prs in 1:length(prs_methods)) {
#     PRS = prs_methods[ind.prs]
# 
#     ## Extract the study number and protein levels
#     currentDF <- as.data.frame(ExpressScan %>% # Extract from ExpressScan instead of the merged ExpressScan_OLINK
#       dplyr::select(.,STUDY_NUMBER, plaquesize_cov, TRAIT) %>%
#       filter(complete.cases(.))) %>%
#       filter_if(~is.numeric(.), all_vars(!is.infinite(.)))
# 
#      # Merge histology data and risk score data on STUDY_NUMBER
#     currentDF <- merge(currentDF, scores, by='STUDY_NUMBER')
#     class(currentDF$Age) = "Numeric"
#   
#     currentDF <- currentDF[, c("Age", "Sex", "ORyear", "CHIP", "PC1", "PC2", PRS, TRAIT, plaquesize_cov)]
#     currentDF <- na.omit(currentDF)
#   
#     # fit for model 2 (includes age + sex)
#     fit <- lm(currentDF[,TRAIT] ~ currentDF[,PRS] + currentDF[,plaquesize_cov] + Age + Sex + ORyear + CHIP + PC1 + PC2, data = currentDF)
#   
#     # fit for model 1 (does not include age + sex)
#     #fit <- lm(currentDF[,TRAIT] ~ currentDF[,PRS] + currentDF[,plaquesize_cov] + ORyear + CHIP + PC1 + PC2, data = currentDF)
#   
#     model_step <- stepAIC(fit, direction = "both", trace = FALSE)
#     print(model_step)
#     print(summary(fit))
#   
#     GLM.results.TEMP <- data.frame(matrix(NA, ncol = 15, nrow = 0))
#     GLM.results.TEMP[1,] = GLM.CON(fit, "ExpressScan_OLINK", PRS, TRAIT, verbose = TRUE)
#     GLM.results = rbind(GLM.results, GLM.results.TEMP)
#   }
# }

### Loop used for protein~prs associations
for (protein in 1:length(olink_proteins_rank)) {
    PROTEIN = olink_proteins_rank[protein]

    for (ind.prs in 1:length(prs_methods)) {
      PRS = prs_methods[ind.prs]

      # Extract the study number and protein levels
      currentDF <- as.data.frame(AEDB_PlasmaProtein_OLINK_rank %>% # Extract from AEDB_PlasmaProtein_OLINK_rank instead of the merged ExpressScan_OLINK
        dplyr::select(.,STUDY_NUMBER, PROTEIN) %>%
        filter(complete.cases(.))) %>%
        filter_if(~is.numeric(.), all_vars(!is.infinite(.)))

      # Merge protein level data and risk score data on STUDY_NUMBER
      currentDF <- merge(currentDF, scores, by='STUDY_NUMBER')
      class(currentDF$Age) = "Numeric"
      
      # Extract phenotypes and covariates from from currentDF
      currentDF <- currentDF[, c("Age", "Sex", "ORyear", "CHIP", "PC1", "PC2", PRS, PROTEIN)]
      currentDF <- na.omit(currentDF)

      ## fit for model WITH age and sex
      fit <- lm(currentDF[,PROTEIN] ~ currentDF[,PRS] + Age + Sex + ORyear + CHIP + PC1 + PC2, data = currentDF)

      ## fit for model WITHOUT age and sex
      #fit <- lm(currentDF[,PROTEIN] ~ currentDF[,PRS] + ORyear + CHIP + PC1 + PC2, data = currentDF)

      model_step <- stepAIC(fit, direction = "both", trace = FALSE)
      print(model_step)
      print(summary(fit))

      GLM.results.TEMP <- data.frame(matrix(NA, ncol = 15, nrow = 0))
      GLM.results.TEMP[1,] = GLM.CON(fit, "ExpressScan_OLINK", PRS, PROTEIN, verbose = TRUE)
      GLM.results = rbind(GLM.results, GLM.results.TEMP)

    }
}

cat("Edit the column names...\n")
colnames(GLM.results) = c("Dataset", "Predictor", "Trait",
                          "Beta", "s.e.m.",
                          "OR", "low95CI", "up95CI",
                          "T-value", "P-value", "r^2", "r^2_adj", "AE_N", "Model_N", "Perc_Miss")

cat("Correct the variable types...\n")
GLM.results$Beta <- as.numeric(GLM.results$Beta)
GLM.results$s.e.m. <- as.numeric(GLM.results$s.e.m.)
GLM.results$OR <- as.numeric(GLM.results$OR)
GLM.results$low95CI <- as.numeric(GLM.results$low95CI)
GLM.results$up95CI <- as.numeric(GLM.results$up95CI)
GLM.results$`T-value` <- as.numeric(GLM.results$`T-value`)
GLM.results$`P-value` <- as.numeric(GLM.results$`P-value`)
GLM.results$`r^2` <- as.numeric(GLM.results$`r^2`)
GLM.results$`r^2_adj` <- as.numeric(GLM.results$`r^2_adj`)
GLM.results$`AE_N` <- as.numeric(GLM.results$`AE_N`)
GLM.results$`Model_N` <- as.numeric(GLM.results$`Model_N`)
GLM.results$`Perc_Miss` <- as.numeric(GLM.results$`Perc_Miss`)
```

```{r CrossSec: Cytokines - linear regression MODEL1 RANK Writing, include=FALSE}
DT::datatable(GLM.results)

# Save the data
cat("Writing results to Excel-file...\n")
### Univariate
library(openxlsx)
write.xlsx(GLM.results,
           file = paste0(OUT_loc, "/",Today,".AEDB_CEA_GLYCC_OLINK_RANK_MODEL1.xlsx"), #change name based on plaque characteristic
           row.names = FALSE, col.names = TRUE, sheetName = "HE")
# Removing intermediates
cat("Removing intermediate files...\n")
rm(TRAIT, currentDF, GLM.results, GLM.results.TEMP, fit, model_step)
```

```{r}
# Calculating PC's for the SlideToolkit plaque characteristics
ExpressScan_plaquecharacteristics <- c("Total_CD34_TissueCount_rank","Total_CD66b_TissueCount_rank","Total_CD68_TissueCount_rank","Total_SMA_TissueCount_rank","Total_HE_TissueCount_rank","Total_EVG_Tissue_rank","Total_GLYCC_Tissue_rank")

ExpressScan_PCs <- ExpressScan[ExpressScan_plaquecharacteristics]

library(MVA)
dt1 <- ExpressScan_PCs
cord <- cor(dt1, use="pairwise.complete.obs")

# Correlation plot
library(corrplot)
corrplot(cord, method="pie")

ExpressScan.pca <- prcomp(cord, retx=TRUE, center=TRUE, scale.=TRUE, tol=NULL, rank.=NULL)
summary(ExpressScan.pca)
print("ExpressScan.pca")
print(ExpressScan.pca)

# Basic screeplot
fviz_eig(ExpressScan.pca)
fviz_eig(ExpressScan.pca, choice = "eigenvalue", addlabels = TRUE)

# Calculating how many PCs are retained
library(factoextra)
eigenvalues <- get_eig(ExpressScan.pca)
eig <- eigenvalues$eigenvalue

library(nFactors)
results <- nCng(eig, details=TRUE)
results

plotuScree(eig, main=paste(results$nFactors,
" factors retained by the CNG procedure",
sep=""))
```

```{r}

# PC's OLINK data + screeplot
## Creating a pairwise correlation matrix
OLINK <-read.csv2("C:\\Users\\anton\\Documents\\UMCU\\Joost_data\\20210506OLINKplasma.csv")
OLINK$X <- NULL

p_missing <- unlist(lapply(OLINK, function(x) sum(is.na(x))))/nrow(OLINK)
sort(p_missing[p_missing > 0], decreasing = TRUE)

OLINKv2 <- OLINK[lapply(OLINK, function(x) sum(is.na(x)) / length(x) ) < 0.70 ]

p_missing <- unlist(lapply(OLINKv2, function(x) sum(is.na(x))))/nrow(OLINKv2)
sort(p_missing[p_missing > 0], decreasing = TRUE)

dt1 <- OLINKv2 [,!colnames(OLINKv2) == "STUDY_NUMBER"]
cord <- cor(dt1, use="pairwise.complete.obs")

# Creating a heatmap and save this
library(pheatmap)
pheatmap(cord, cluster_row = TRUE, cluster_cols = TRUE, fontsize = 2, filename = "20211105 Heatmap OLINK.pdf")


# Calculating the PC's
olink.pca <- prcomp(cord, retx=TRUE, center=TRUE, scale.=TRUE, tol=NULL, rank.=NULL)
summary(olink.pca)

fviz_eig(olink.pca)
fviz_eig(olink.pca, choice = "eigenvalue", addlabels = TRUE)


# Determining how many PC's have to be retained
eigenvalues <- get_eig(olink.pca)
eig <- eigenvalues$eigenvalue
results <- nCng(eig, details=TRUE)
results

plotuScree(eig, main=paste(results$nFactors,
" factors retained by the CNG procedure",
sep=""))

#nCng(cord, cor=TRUE, model = "components", details=TRUE)

```
